Modular Inverse Reinforcement Learning on Human Motion

### 2015-02-11 paper.tex
experiment section

Different weights can yield different performance. Let $w_1, w_2, w_3$ be
weights for the task of target collection, obstacle avoidance, and path
following, respectively. Let $w$ be the vector of $(w_1, w_2, w_3)$. An agent
with $w = (1, 0, 0)$ only collect targets, and one with $w = (0, 0.5, 0.5)$ may
avoid the obstacles and follow the path.

To obtain the weights given the samples, we need to use the Inverse Modular
Reinforcement Learning technique \cite{rothkopf2013modular}.

@@

\begin{figure}[h!]
\centering
\includegraphics[width=0.24\textwidth]{task_1.png}
\includegraphics[width=0.24\textwidth]{task_2.png}
\includegraphics[width=0.24\textwidth]{task_3.png}
\includegraphics[width=0.24\textwidth]{task_4.png}
\caption{}
\label{fig:exp}
\end{figure}

From a different perspective, can we find a weight vector to best interpret
human's behavior. In Figure~\ref{fig:exp}, same as Figure~\ref{fig:avatar}, the
red circles are obstacles. The blue circles are targets. The gray line is the
path. The black lines are trajectories of human.
%TODO

We make some constraints on our learning agent to make it walk like a human.
We can find in the human trajectories that humans walk smoothly. They don't turn
sharply.  Our agent is allowed to do three actions --- going straight ahead,
turning left with a small step, and turning right with a small step.

To make our weights represent the significance of the modules, we
normalize the sub-MDPs with the unit (positive or negative) rewards. The reward
is 1 for collecting a target, -1 for running into an obstacle. We define the
value function directly for the path module to have a path following performance.

The agent only considers the closest target and the closest obstacle.

We assume that our learning agent only knows the three sub-MDPs and the human
data. It looks at the human behavior, and finds the weights that can interpret
such behavior. 
Using such weights, the trajectories of our agents are drawn in the green lines.
We can tell that in the left figure, the agent puts a large weight on
path-following.  In the right figure, it puts weights on all sub-MDPs.

\begin{figure}[h!]
\centering
\includegraphics[width=0.4\textwidth]{objValuesTask1.png}
\includegraphics[width=0.4\textwidth]{objValuesTask2.png}
\includegraphics[width=0.4\textwidth]{objValuesTask3.png}
\includegraphics[width=0.4\textwidth]{objValuesTask4.png}
\caption{Heatmaps of the objective values of different weights for the four
tasks, respectively. The red zones indicate higher values. The upper two are
Task 1 and 2. The lower two are Task 3 and 4.}
\label{fig:heatmap}
\end{figure}

\begin{table}[h]
\centering
\begin{tabular}{| l| l| l |}
\hline
Average by Task & Num Targs Hit & Num Obst Hit\\
\hline
1 & 2.34 &  2.13\\
\hline
2 & 3.03 &  \bf{0.13}\\
\hline
3 & \bf{10.19} & 2.28\\
\hline
4 & \bf{9.88} &  \bf{0.03}\\
\hline
\end{tabular}
\caption{Number of targets hit and number of obstacles hit of the humans.}
\end{table}

\begin{table}[h]
\centering
\begin{tabular}{| l| l| l |}
\hline
Average by Task & Num Targs Hit & Num Obst Hit\\
\hline
1 & 1.25 & 1.62\\
\hline
2 & 3.62 & \bf{2.37}\\
\hline
3 & \bf{5.14} & 3.14\\
\hline
4 & \bf{5.00} & \bf{2.00}\\
\hline
\end{tabular}
\caption{Number of targets hit and number of obstacles hit of the learning agent.}
\end{table}

In Figure~\ref{fig:heatmap}, we show the object values for different weights.

@@

Weighted sum of Q function is one way to combine multiple sub-MDPs. We also
propose other ways including, for example, scheduling between different modules,
with only one active at one time. This is also called skilled in the literature
\cite{konidaris2009skill}. However, we adopt the weighted sum approach
because this is more reasonable for human behavior. When a human tries to collect
targets while avoiding obstacles, these two modules are expected to be both
active. A scheduling approach may yield frequent oscillation between these two
modules.

>>>>

Different weights can yield different performance. Let $w_1, w_2, w_3$ be
weights for the task of target collection, obstacle avoidance, and path
following, respectively. Let $w$ be the vector of $(w_1, w_2, w_3)$. An agent
with $w = (1, 0, 0)$ only collects targets, and one with $w = (0, 0.5, 0.5)$ may
avoid the obstacles and follow the path.

To obtain the weights given the samples, we need to use the Inverse Modular
Reinforcement Learning technique \cite{rothkopf2013modular}. We use a maximum
likelihood method here.
\begin{equation}
\label{eq:irl}
\max_w \prod_t \frac{e^{\eta Q(s^{(t)}, a^{(t)})}}{\sum_b e^{\eta Q(s^{(t)}, b)}}
\end{equation}
where $s^{(t)}$ is the state at time $t$, and $a^{(t)}$ is the action at time
$t$, which are both from samples. $Q(s, a) = \sum_i w_i Q_i(s_i, a)$, as defined
before. $\eta$ is a hyperparameter that determines the consistency of human's
behavior. The larger $\eta$ is, the algorithm more tends to overfit the data.

The intuition of Equation~\ref{eq:irl} is that if an action is observed from the
sample, then the Q value of taking that action should be larger compared to Q
values of taking other actions.

@@

We train the modules first to get $Q_i$ before running the inverse reinforcement
learing algorithm. For each module, the agent only considers the closest target
and the closest obstacle. For the path module, the path is defined as segments
of waypoints, so the closest waypoint is considered. The agent takes the
distance and angle to the closest objects as the state representation.

To make our weights represent the significance of the modules, we
normalize the sub-MDPs with the unit (positive or negative) rewards. The reward
is 1 for collecting a target, -1 for running into an obstacle. We define the
value function directly for the path module to have a path following
performance, as it is tricky to give reward for such performance.

We make some constraints on our learning agent to make it walk like a human.
We can find in the human trajectories that humans walk smoothly. They don't turn
sharply.  Our agent is only allowed to do three actions --- going straight ahead,
turning left with a small step, and turning right with a small step.

\begin{figure}[h!]
\centering
\includegraphics[width=0.24\textwidth]{task_1.png}
\includegraphics[width=0.24\textwidth]{task_2.png}
\includegraphics[width=0.24\textwidth]{task_3.png}
\includegraphics[width=0.24\textwidth]{task_4.png}
\caption{The trajectories of humans and our agent in four different tasks. The
upper two are Task 1 and 2. The lower two are Task 3 and 4.}
\label{fig:exp}
\end{figure}

\begin{table}[h!]
\centering
\begin{tabular}{| l| l| l |}
\hline
Average by Task & Num Targs Hit & Num Obst Hit\\
\hline
1 & 2.34 &  2.13\\
\hline
2 & 3.03 &  \bf{0.13}\\
\hline
3 & \bf{10.19} & 2.28\\
\hline
4 & \bf{9.88} &  \bf{0.03}\\
\hline
\end{tabular}
\caption{Number of targets hit and number of obstacles hit of the humans.}
\label{tb:human}
\end{table}

\begin{table}[h!]
\centering
\begin{tabular}{| l| l| l |}
\hline
Average by Task & Num Targs Hit & Num Obst Hit\\
\hline
1 & 1.25 & 1.62\\
\hline
2 & 3.62 & \bf{2.37}\\
\hline
3 & \bf{5.14} & 3.14\\
\hline
4 & \bf{5.00} & \bf{2.00}\\
\hline
\end{tabular}
\caption{Number of targets hit and number of obstacles hit of the learning agent.}
\label{tb:agent}
\end{table}

We report the results in Figure~\ref{fig:exp}, same as Figure~\ref{fig:avatar},
the red circles are obstacles. The blue circles are targets. The gray line is
the path. The black lines are trajectories of human --- each line represents one
human trajectory.
Using the weights derived from the algorithm, the trajectories of our agents are
drawn in the green lines.

Although humans don't agree each other on how to avoid obstacles or collecting
targets, our agent can figure out what the humans are doing, and perform the
similar performance. Table~\ref{tb:human} and Table~\ref{tb:agent} show the
number of targets hit and number of obstacles hit. Note that the numbers bolded
are active modules in the corresponding task. We can observe that humans still
do better than our agent in these tasks.

\begin{figure}[h!]
\centering
\includegraphics[width=0.4\textwidth]{objValuesTask1.png}
\includegraphics[width=0.4\textwidth]{objValuesTask2.png}
\includegraphics[width=0.4\textwidth]{objValuesTask3.png}
\includegraphics[width=0.4\textwidth]{objValuesTask4.png}
\caption{Heatmaps of the $\log$ of values of Equation~\ref{eq:irl} for different
weights for the four tasks, respectively. The red zones indicate higher values.
From left to right are Task 1 to Task 4.}
\label{fig:heatmap}
\end{figure}

In Figure~\ref{fig:heatmap}, we show the $\log$ of values of Equation~\ref{eq:irl} for
different weights. We can observe the centroids of red zones move for different
tasks. It stays at the origin in Task 1, so none of target and obstacle modules
are active. It moves away from the origin when a module is active.

@@

Weighted sum of Q functions is one way to combine multiple sub-MDPs. We also
propose other ways including, for example, scheduling between different modules,
with only one active at one time. This is also called skilled in the literature
\cite{konidaris2009skill}. However, we adopt the weighted sum approach
because this is more reasonable for human behavior. When a human tries to collect
targets while avoiding obstacles, these two modules are expected to be both
active. A scheduling approach may yield frequent oscillation between these two
modules.
